---
title: "Movie recommendation engine"
output: 
  html_document: 
    smart: no
---

https://github.com/jeknov/movieRec/blob/master/movieRec.R
https://muffynomster.wordpress.com/2015/06/07/building-a-movie-recommendation-engine-with-r/
https://www.analyticsvidhya.com/blog/2016/03/practical-guide-principal-component-analysis-python/
http://blog.nycdatascience.com/student-works/machine-learning/movie-rating-prediction/


### Project details: 


#### Domain: Entertainment
#### Project: IMDB - Recommendation
#### Data: movie_metadata.csv

#### Description
- This dataset contains 28 variables for 5043 movies, spanning across 100 years in 66 countries. There are 2399 unique director names, and thousands of actors/actresses. Below are the 28 variables:

movie_title                                          
color                                                          
num_critic_for_reviews                                               
movie_facebook_likes                                                
duration                                                        
director_name                                                       
director_facebook_likes                                                       
actor_3_name                                                              
actor_3_facebook_likes                                                      
actor_2_name                                                                           
actor_2_facebook_likes                                                                
actor_1_name                                                                         
actor_1_facebook_likes                                                        
gross                                                                
genres                                                                       
num_voted_users                                                                  
cast_total_facebook_likes                                                            
facenumber_in_poster                                                                  
plot_keywords                                                                   
movie_imdb_link                                                                 
num_user_for_reviews                                                          
language                                                                  
country                                                                  
content_rating                                                               
budget                                                                                 
title_year                                                                      
imdb_score                                                                          
aspect_ratio                                                                

#### Objective
- Given that thousands of movies were produced each year, is there a better way for us to tell the greatness of movie without relying on critics or our own instincts? And try to create a recommendation engine with this dataset.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Loading required packages

```{r}

library(tidyr)
library(dplyr)
library(reshape2)
library(ggplot2)

```


### Loading the data

```{r}

movies <- read.csv('movie_metadata.csv',stringsAsFactors=FALSE)
glimpse(movies)
summary(movies)

```

The missing values are present in:

num_critic_for_reviews(50)                                
duration(15)                                      
director_facebbok_likes(104)                       
actor_3_facebook_likes(23)                       
actor_1_facebook_likes(7)                             
gross(884)                               
facenumber_in_poster(13)                              
num_user_for_reviews(21)                               
budget(492)                                        
title_year(108)                                          
actor_2_facebook_likes(13)                          
aspect_ratio(329)                                     


### Removing the duplcate movies

```{r}

idx_dup <- which(duplicated(movies$movie_title)==TRUE)
movies <- movies[-idx_dup,]


```


### Finding columns and rows with missing values using function
```{r}
miss <- function(x) {
 cat('\nThe columns having missing values are:')
 for(i in 1:ncol(x)) {
  if(length(x[i][is.na(x[i])]) > 0) {
    cat('\n',names(x[i]),':',length(x[i][is.na(x[i])]))
  }
 } 
  cat('\n\nThe number of rows with missing values are:',nrow(x)-nrow(na.omit(x)))
  cat('\nThe proportion of missing values is:',signif((nrow(x)-nrow(na.omit(x)))/nrow(x),2))
}

miss(movies)
```


### Removing rows with missing values in budget and gross
### Missing values in rest of the columns can be replaced with zeros

```{r}

# remove the rows having missing values in budget and gross
idx <- which(is.na(movies$budget)==TRUE)
movies <- movies[-idx,]

idx <- which(is.na(movies$gross)==TRUE)
movies <- movies[-idx,]
 
# replace with zeros
movies[is.na(movies)] <- 0

# checking the missing values
miss(movies)
 

```

Now the dataset is clean and ready


### Write into csv file so as to do visualizations in tableau
### check the Visualizations here:

https://public.tableau.com/profile/bharadwaj.tadikonda#!/vizhome/Imdb_recommendation/ImdbscorevsNumvotedUsers

```{r}

write.csv(movies,'movies_tab.csv')


```


Can infer the following from plots:

- Majority of the scores are in range of 6-8

- Countries like China, Hong kong, India, Italy, Japan, Mexico, Spain, UK have produced the average score of above 7 (min. count of movies is 10)

- UK produced lots of movies (313) with average Imdb score of 7

- US produced largest number of movies(2993), however their average imdb score is 6.5

- We can find that over the years(spcifically after 1975) the quality of movies(as per score) has come down. This might be partly due to the fact that large number of movies are being made.

- As the "movie facebook likes" increases, the score also increases. There are few outliers too like for instance "Interstellar" got the maximum likes(>350k) but score is only 8.6 where as "Godfather" got less than 50k likes but its score is 9.2

- There is positive relationship between "cast total FB likes" and Imdb score, however its not as strong as that of "movie facebook likes"

- As the duration of movie increases, the score also increases

- Generally speaking as the score of movie increases, the gross also increases. However the budgets for high gross earning movies is also high. For instance, movie like "Avatar" which made whopping gross had gross profit margin of 69% where as "A separation" made on comparitively less budget made gross profit margin of 93%

- The most succesful director(with min. 10 movies) is steven spielberg (both in terms of score and gross profit margin %)

- The Top actor score wise(with min. 10 movies) are Tom hanks, Leonardo Dicaprio and Clint Eastwood

- The Top actor-director combination(with min. 3 movies together) are Christain Bale-Christopher Nolan, Robert De Niro -Martin Scorsese, Leonardo Dicaprio -Martin Scorsese, Michael Biehn-James Cameron, Michael Moore-Michael Moore.

- More the "Num Voted users", better is the score. The movie with highest votes is "The shawshank Redemption"(9.3), followed by "The Dark Knight"(9) and "Inception"(8.8)


### Correlation analysis

```{r}

library(corrplot)
x <- cor(movies[,unlist(lapply(movies, is.numeric))])

#corrplot(x,method='number',number.cex=0.75,col='blue')
corrplot(x,method='number',col='black',cl.pos = "n",number.cex=0.75)

xdf <- as.data.frame(as.table(x))
xdf <- xdf %>% filter(abs(Freq) > 0.5 & Freq != 1) %>% arrange(desc(Freq))
 
View(xdf)

```

Can infer that:

<For imdb score>

- The "imdb_score" has positive correlation with "num_voted_users". More the voted users, movies tend to have high rating. 

- The "imdb_score" has small but positive correlation with "duration". Long movies tend to have high rating.

- The "imdb_score" has very small but positive correlation with the "director_facebook_likes", meaning a popular director does not necessarily mean his directed movie is great.

- The "imdb_score" has very small but positive correlation with the "actor_1_facebook_likes", meaning that an leading actor is popular in 
social network does not mean that a movie is high rating

- The "imdb_score" has small but negative correlation with "facenumber_in_poster". It is perhaps not a good idea to have many faces in movie poster if a movie wants to be great.

- The "imdb_score" has almost no correlation with "budget". Throwing money at a movie will not necessarily make it great.

<For others>

-The "cast_total_facebook_likes" has a strong positive correlation with the "actor_1_facebook_likes", and has smaller positive correlation with both "actor_2_facebook_likes" and "actor_3_facebook_likes"

- The "num_user_for_reviews" & "num_critic_for_reviews"" has strong positive correlation with the "num_voted_users", meaning the users who write reviews and are critics also vote for the movie

- The "movie_facebook_likes" has strong correlation with "num_critic_for_reviews", meaning that the popularity of a movie in social network can be largely affected by the critics

- The movie "gross" has strong positive correlation with the "num_voted_users"

- The "movie_facebook_likes" has relatively large correlation with the "num_voted_users"

### Correlation implies linear relationship. Let's validate it through scatter plot
### Multicollinearity should not exist for linear regression

```{r}
ggplot(movies,aes(actor_1_facebook_likes,cast_total_facebook_likes))+geom_point()+geom_smooth(method = 'lm')
ggplot(movies,aes(num_user_for_reviews,num_voted_users))+geom_point()+geom_smooth(method = 'lm')
ggplot(movies,aes(movie_facebook_likes,num_critic_for_reviews))+geom_point()+geom_smooth(method = 'lm')
ggplot(movies,aes(num_voted_users,gross))+geom_point()+geom_smooth(method = 'lm')


```



### Selecting the variables for linear regression


```{r}

#remove variables that correspond to post release
  #num_critic_for_reviews 
  #gross
  #num_voted_users
  #num_user_for_reviews
  #movie_facebook_likes

#remove variables having high collinearity
  #cast_total_facebook_likes
  
#remove other variables 
  #genres
  #plot_keywords
  #movie_imdb_link
  #director_name
  #actor_1_name
  #actor_2_name
  #actor_3_name
  #movie_title
  #country

```

### Split the data into train and test


```{r}

# removing the variables 

#movies_sub <- movies %>% select(-c(num_critic_for_reviews, gross, num_voted_users, num_user_for_reviews, movie_facebook_likes, cast_total_facebook_likes, genres, plot_keywords, movie_imdb_link,director_name,actor_1_name,actor_2_name,actor_3_name,movie_title,country,aspect_ratio))

#divide the new data

library(caret)

set.seed(100)
idx <- createDataPartition(movies$imdb_score,p=0.7,list = FALSE)

movies_train <- movies[idx,]
movies_test <- movies[-idx,]

#standardise the data
movies_train1 <- movies_train
movies_test1 <- movies_test

movies_train1[,c(3:6,8:9,13:14,19,23,25,28)] <- scale(movies_train1[,c(3:6,8:9,13:14,19,23,25,28)], center = TRUE, scale = TRUE)

movies_test1[,c(3:6,8:9,13:14,19,23,25,28)] <- scale(movies_test1[,c(3:6,8:9,13:14,19,23,25,28)], center = TRUE, scale = TRUE)

```


### Defining parameters

```{r}

#Define the training control

fitcontrol <- trainControl(method="cv",number=10,savePredictions='final')
#fitcontrol <- trainControl(savePredictions='final')

#Defind predictors and outcome

#predictors <- c("duration","director_facebook_likes","actor_3_facebook_likes","actor_1_facebook_likes","facenumber_in_poster","content_rating","budget","title_year","actor_2_facebook_likes","language","color")
predictors <- c("duration","director_facebook_likes","actor_3_facebook_likes","actor_1_facebook_likes","facenumber_in_poster","budget","title_year","actor_2_facebook_likes")


outcome <- "imdb_score"

```


### Initializing the results table

```{r}

prediction <- data.frame()

```

### Creating error measures(RMSE, Rsquared) to validate the model
### RMSE    : Root Mean squared Error. It tells the error in prediction
### Rsquared: It tells the poportion of variance explained by the model

```{r}

Error_measures <- function(actual,predicted)
{
  SSE <- sum((actual-predicted)^2)
  SST <- sum((actual-mean(actual))^2)
  
  R2 <<- 1-(SSE/SST)
  
  error <- actual-predicted
  RMSE <<- sqrt(mean(error^2))
}

```



### Linear Regression Training


```{r}

set.seed(101)
model_lm <- train(movies_train1[,predictors],
                  movies_train1[,outcome],
                  method = 'lm',
                  #preProcess = c('center','scale'),
                  tuneLength = 9,
                  trControl = fitcontrol)

model_lm



```


### Validating Linear Regression Assumptions

```{r}

#Linearity & Homoscedasticity

residuals <- model_lm$pred[,'obs']-model_lm$pred[,'pred']
plot(model_lm$pred[,'pred'],residuals,xlab='Fitted values',main="Fitted vs Residuals")

#Normality

qqnorm(residuals)
qqline(residuals)

#Multicollinearity
library(car)
vif(model_lm$finalModel)

```

Can infer that:

- Linearity & homoscedasticity exists(though not perfect)
- Normality exists(though slight deviation)
- VIF<=4, so we can say multicollinearity doesnot exist between the predictor variables
- Since all these assumptions hold good, we can proceed with linear regression model


### Linear Regression Validation


```{r}

#prediction
predict_model_lm <- predict(model_lm,movies_test1[,predictors])

#calculating error measures
Error_measures(movies_test1$imdb_score,predict_model_lm)

#Building the result into table
prediction <<- prediction[-1,]
temp1 <- data.frame(RMSE=RMSE,Rsquared=R2)
prediction <<- rbind(prediction, lin_reg=temp1 )
prediction['lin_reg',]

```




### Decision Tree


```{r}

set.seed(101)
model_rpart <- train(movies_train1[,predictors],
                  movies_train1[,outcome],
                  method = 'rpart',
                  #preProcess = c('center','scale'),
                  tuneLength = 9,
                  trControl = fitcontrol)

model_rpart

```

### Decision Tree Validation


```{r}

#prediction
predict_model_rpart <- predict(model_rpart,movies_test1[,predictors])

#calculating error measures
Error_measures(movies_test1$imdb_score,predict_model_rpart)

#Building the result into table
prediction <<- prediction[-2,]
temp1 <- data.frame(RMSE=RMSE,Rsquared=R2)
prediction <<- rbind(prediction, decision_tree=temp1 )
prediction['decision_tree',]

#plot the tree
plot(model_rpart$finalModel)
text(model_rpart$finalModel)

```

### random forest


```{r}

set.seed(101)
model_rf <- train(movies_train1[,predictors],
                  movies_train1[,outcome],
                  method = 'rf',
                  #preProcess = 'scale',
                  importance = TRUE,
                  tuneLength = 7,
                  trControl = fitcontrol)

model_rf

```

### random forest Validation


```{r}

#prediction
predict_model_rf <- predict(model_rf,movies_test1[,predictors])

#calculating error measures
Error_measures(movies_test1$imdb_score,predict_model_rf)

#Building the result into table
prediction <<- prediction[-3,]
temp1 <- data.frame(RMSE=RMSE,Rsquared=R2)
prediction <<- rbind(prediction, random_forest=temp1 )
prediction['random_forest',]

```

### Plot the error in random forest model

```{r}

#plot the error
plot(model_rf$finalModel)

```

Can infer that:

- Initially there is lot of volatility but the curve stabilizes after aound 50 trees 

### Variable importance


```{r}


#check the variables important to the model

#varImp(model_rf$finalModel)
varImpPlot(model_rf$finalModel,main="Variable Importance")

```

Can infer that:

- The most important factor affecting movie rating is duration. The longer the movie, the better is the rating
- Budget is also important, although we have seen earlier that budget has got very less correlation with rating
- Facebook popularity of director is very important to the rating.
- Facebook popularity of Top 3 actors also effects the rating.
- Title year also has little impact on the rating.
- Facenumber in poster is having very less effect on the rating




